The Mass Distribution of Stellar-Mass Black Holes

### 2010-10-22 paper.tex
Made Vicky's brief changes.

\begin{abstract}
  We perform a Bayesian analysis of the mass distribution of
  stellar-mass black holes using the observed masses of 15 low-mass
  X-ray binary systems undergoing Roche lobe overflow and five
  higher-mass, wind-fed X-ray binary systems.  Modeling the mass
  distribution both parametrically---as a power law, exponential,
  gaussian, combination of two gaussians, or log-normal
  distribution---and non-parametrically---as histograms with varying
  numbers of bins---we explore the distribution in the context of
  Markov Chain Monte Carlo calculations of model parameters.  We give
  confidence bounds on the shape of the mass distribution in the
  context of each model and compare the models with each other by
  calculating their Bayesian evidence.  The mass distribution of the
  low-mass systems is best fit by a power-law, while the distribution
  of the complete sample is best fit by the exponential and power-law
  models (but with significantly different parameters than the
  lower-mass sample).  We address the existence of a ``gap'' between
  the most massive neutron stars and least massive black holes,
  finding that the best model (the power law) fitted to the low-mass
  systems only gives a minimum black-hole mass of 4.3 $\Msun$ (90\%
  confidence), while the best models (the exponential and power-law)
  fitted to all 20 systems gives a minumum black-hole mass of 4.5 and
  5.2 $\Msun$ (90\% confidence), respectively.  We conclude that our
  sample of black hole masses provides strong evidence of a gap
  between the maximum neutron star mass and minimum black hole mass.
\end{abstract}

@@

The mass function defines a lower limit on the mass: $f(M) < M$.  To
accurately determine the mass of the black hole, the inclination $i$
and mass ratio $q$ must be measured.  Ideally, this can be
accomplished by fitting ellipsoidal light curves and study of the
rotational broadening of spectral lines from the secondary, but even
in the ideal case (see, e.g., \citet{Cantrell2010} on A0620) this
procedure is complicated.  In particular, contributions from an
accretion disk and hot spots in the disk can significantly distort the
measured inclination and mass ratios.  For some systems (e.g.\ GS 1354
\citep{Casares2009}) strong variability prevents determination of the
inclination from the lightcurve at all; in these cases an upper limit
on the inclination often comes from the observed lack of eclipses in
the lightcurve.  In general, accurately determining $q$ and $i$
requires a careful system-by-system analysis.

@@

In Figure \ref{fig:masses} we show the combination of these individual
mass distributions into one single distribution.  Note that this
combination is not a good estimator of the true mass distribution of
black hole masses.  To understand why this is not a good estimator of
the underlying mass distribution, suppose that the observations
implied an identical mass distribution for each object.  Then the
combined distribution is equal to the individual distributions, but in
fact we can be confident that each of the objects was drawn from a
distribution that was narrower by a factor of approximately
$\sqrt{N}$.  \citet{Mandel2010} discusses this issue in detail.  We
include Figure \ref{fig:masses} here only to give a sense of the range
of masses of black holes in our sample; we will come to estimators for
the mass distribution.

\begin{figure}
  \plottwo{plots/masses}{plots/masses-high}

  \caption{\label{fig:masses} The mass distributions in Figure
    \ref{fig:low-masses} (left) and both \ref{fig:low-masses} and
    \ref{fig:high-masses} (right) combined into a single
    distribution. Note that this combination is not a good estimator
    of the true mass distribution of black hole masses in the
    galaxy---see, e.g., \citet{Mandel2010}.  We include it here only
    to give a sense of the range and general shape of the mass
    distribution of black holes in our sample.}
\end{figure}

@@

In this section we apply a Bayesian analysis to various models for the
underlying mass distribution from which the low-mass systems in Table
\ref{tab:sources} were drawn.  We will treat the combined low- and
higher-mass systems in Section \ref{sec:higher-mass}.  The end result
will be the probability distribution for the parameters of each model
implied by the data from Section \ref{sec:systems} in combination with
our prior assumptions about the probability distribution for the
parameters.  Bayes' rule relates these quantities.  For a model with
parameters $\vtheta$ in the presence of data $d$, Bayes' rule states
\begin{equation}
  \label{eq:Bayes-rule}
  p(\vtheta | d) = \frac{p(d | \vtheta) p(\vtheta)}{p(d)}.
\end{equation}
Here, $p(\vtheta|d)$, called the posterior probability distribution
function, is the probability distribution for the parameters $\vtheta$
implied by the data $d$; $p(d|\vtheta)$, called the likelihood, is the
probability of observing data $d$ given that the model parameters are
$\vtheta$; $p(\vtheta)$, called the prior, reflects our estimate of the
probability of the various model parameters in the absence of any
data; and $p(d)$, called the evidence, is an overall normalizing
constant ensuring that
\begin{equation}
  \int d\theta\, p(\vtheta|d) = 1,
\end{equation}
whence
\begin{equation}
  \label{eq:evidence-def}
  p(d) = \int d\vtheta\, p(d|\vtheta) p(\vtheta).
\end{equation}

@@

Our calculation of the likelihood of each observation does not include
any attempt to account for selection effects in the observations.  We
simply assume (almost certainly incorrectly) that any black hole drawn
from the underlying mass distribution is equally likely to be
observed, leaving the analysis of selection effects to future work.
(During the preparation of this work, we became aware of
\citet{Ozel2010}, which finds that selection effects do not greatly
affect an analysis similar to ours.)

@@

The marginalized distribution for $\alpha$ is broad, with
\begin{equation}
  -11.8 < \alpha < 6.8
\end{equation}
enclosing 90\% of the probability (excluding 5\% on each side).  We
have $p(\alpha < 0) = 0.6$.  The median value is $\alpha = -3.35$.
The broadness of the marginalized distribution for $\alpha$ comes from
the need to match the relatively narrow peak in the mass distribution
apparent in Figure \ref{fig:masses}.  When $\alpha$ is negative, the
resulting mass distribution slopes down; $\Mmin$ is constrained to be
near the lowest mass of the observed black holes, while $\Mmax$ is
essentially irrelevant.  Conversely, when $\alpha$ is positive and the
mass distribution slopes up, $\Mmax$ must be close to the largest mass
observed, while $\Mmin$ is essentially irrelevant.  Figure
\ref{fig:power-law-2D} illustrates this effect, showing the
correlations between $\alpha$ and $\Mmin$ and $\alpha$ and $\Mmax$.
When we include the higher-mass systems in the analysis, the long tail
will eliminate this effect by bringing both $\Mmin$ and $\Mmax$ into
play for all values of $\alpha$.

@@

Figure \ref{fig:two-gaussian} shows the marginalized distributions for
the two-Gaussian model parameters from our MCMC runs.  We find $\alpha
> 0.8$ with 62\% probability, clearly favoring the first Gaussian.
The distribution for $\mu_1$ and $\sigma_1$ are similar to those of
the single Gaussian displayed in Figure \ref{fig:gaussian}, indicating
that the first Gaussian has found the peak mass samples in Figure
\ref{fig:masses}.  The second Gaussian's parameter distributions are
much broader.  The second Gaussian appears to be sampling the tail of
the mass samples (see Figure \ref{fig:masses}).  In spite of the
ability to match a more complicated distribution, we find that this
model is strongly disfavored relative to the single-Gaussian model for
this dataset: $p(\textnormal{Gaussian}|d) / p(\textnormal{Two
  Gaussian}|d) \simeq 4.7$ (see Section \ref{sec:model-selection} for
discussion).

\begin{figure}
  \begin{center}
    \plotone{plots/two-gaussian}
  \end{center}
  \caption{\label{fig:two-gaussian} The marginal distributions for the
    five parameters of the two-Gaussian model.  We have $\alpha > 0.8$
    with 62\% probability, favoring the first of the two Gaussians.
    The distributions for $\mu_1$ and $\sigma_1$ are similar to those
    of the single Gaussian model displayed in Figure
    \ref{fig:gaussian}; the second Gaussian's parameter distributions
    are much broader (recall that we constrain $\mu_2 > \mu_1$).  The
    second Gaussian is attempting to fit the tail of the mass samples
    (see Figure \ref{fig:masses}).  The extra degrees of freedom in
    the distribution from the second Gaussian do not provide enough
    extra fitting power to compensate for the increase in parmeter
    space, however: the two-Gaussian model is disfavored relative to
    the single Gaussian by a factor of $4.7$ on this dataset (see
    Section \ref{sec:model-selection} for discussion).}
\end{figure}

@@

There are two broad ways to compare models in the context of the MCMC
analysis of Section \ref{sec:models}.  The first way, which we discuss
in Section \ref{sec:direct-evidence}, is to use the MCMC samples
from the various models to estimate the single model evidence,
$p(d|M_i)$, appearing in Equation \eqref{eq:model-evidence-def}.  The
second way is to treat the model literally as a parameter in a second
MCMC analysis.  In the second MCMC analysis, one proposes jumps not
between different $\vtheta_i$ within a particular model, but from
parameters $\vtheta_i$ in model $M_i$ to parameters $\vtheta_j$ in
$M_j$.  The evidence for model $M_i$ is proportional to the number of
times this MCMC has a sample in the $M_i$ parameter space.  This
technique is called a ``reversible-jump MCMC'' \citep{Green1995}; we
discuss it in Section \ref{sec:reversible-jump-mcmc}.

\subsection{Direct Evidence Calculations}
\label{sec:direct-evidence}

To compute $p(M_i|d)$ from Equation \eqref{eq:model-evidence-def}
using the MCMC samples we discussed in Section \ref{sec:models} we
need a way to compute the single-model evidence, $p(d|M_i)$.  The
remaining terms in Equation \eqref{eq:model-evidence-def} are either
under our control (the prior $p(M_i)$) or a normalization factor that
we can compute using Equation \eqref{eq:multi-model-evidence-def} once
we have the single model evidence (the evidence $p(d)$).  Begin by
re-writing Equation \eqref{eq:bayes-explicit-model}:
\begin{equation}
  \frac{p(\theta_i|d,M_i)}{p(d|\theta_i,M_i)} = \frac{p(\theta_i|M_i)}{p(d|M_i)}.
\end{equation}
Integrating both sides over all $\theta_i$, recalling that the prior
is normalized, we have 
\begin{equation}
  \int d\theta_i\, \frac{p(\theta_i|d,M_i)}{p(d|\theta_i,M_i)} = \frac{1}{p(d|M_i)}.
\end{equation}
We can estimate the integral using our MCMC samples (which are
distributed with local density $p(\theta_i|d,M_i)d\theta_i$) via
\begin{equation}
  \int d\theta_i\, \frac{p(\theta_i|d,M_i)}{p(d|\theta_i,M_i)} \approx
  \frac{1}{N_i} \sum_j \frac{1}{p(d|\theta_{ij}, M_i)},
\end{equation}
where the $j$th MCMC parameter sample (from a total $N_i$) is
$\theta_{ij}$.  Thus we have
\begin{equation}
  \frac{1}{p(d|M_i)} \approx \frac{1}{N_i} \sum_j \frac{1}{p(d|\theta_{ij}, M_i)}.
\end{equation}
This technique is known as the ``harmonic mean approximation'' for the
single-model evidence $p(d|M_i)$.

\citet{Weinberg2010} discusses techniques to estimate the integral
over parameter space in Equation \eqref{eq:model-evidence-def} by
adaptively partitioning the parameter volume into non-overlapping
boxes surrounding the points in the MCMC sample.  We implemented the
algorithms from that paper, but found them, in this case, to give
results that depended strongly on the parameters in the algorithm and
were sometimes inconsistent with both the harmonic mean calculations
in this Section and the reversible jump calculations in Section
\ref{sec:reversible-jump-mcmc}.  Since the harmonic mean and
reversible jump calculations are in agreement with each other, we will
not discuss these direct integration techniques further.

The harmonic mean estimate of the evidence is known for having large
statistical uncertainty due to the tendency for a few low-likelihood
points to dominate the sum \citep{Weinberg2010}.  Therefore, it is
useful to have multiple independent estimates of the evidence for a
given model.  In Section \ref{sec:reversible-jump-mcmc}, we will
discuss the reversible jump algorithm for estimating the evidence for
our models; the results presented here are in broad agreement with the
results in that Section.  In Figure \ref{fig:harmonic-evidence} we
display the harmonic mean estimator for the evidence for the various
models discussed in Section \ref{sec:models}.  In addition, we have
performed 10000 calculations of the harmonic mean estimator using
``mock'' MCMC data drawn randomly from our true MCMC with replacement.
The errorbars in Figure \ref{fig:harmonic-evidence} indicate the 10\%
and 90\% quantiles from this bootstrap calculation.  Recall that we
assume that no model is a priori more likely than any other, so the
model priors $p(M_i)$ are equal.

\begin{figure}
  \begin{center}
    \plotone{plots/evidence}
  \end{center}
  \caption{\label{fig:harmonic-evidence} Harmonic mean evidence for
    the various models from Section \ref{sec:models} (see Section
    \ref{sec:direct-evidence}).  In increasing order along the
    $x$-axis, the models are the power-law of Section
    \ref{sec:power-law} (PL), the decaying exponential of Section
    \ref{sec:exponential} (E), the single Gaussian of Section
    \ref{sec:gaussian} (G), the double Gaussian of Section
    \ref{sec:gaussian} (TG), and the one-, two-, three-, four-, and
    five-bin histogram models of Section
    \ref{sec:non-parametric-models} (H1, H2, H3, H4, H5,
    respectively).  The errorbars are bootstrap estimates of the 10\%
    and 90\% bounds on the harmonic mean estimator.  The power-law and
    Gaussian are the favored models, with the single-bin histogram
    coming third (but the 10-90\% bootstrap range for the one-bin
    histogram includes the possibility of being the most favored
    model).  This is not surprising, given the single-peaked nature of
    the lower-mass sample.}
\end{figure}

The results of the harmonic mean calculation indicate that the
power-law and Gaussian models are most favored, with the single-bin
histogram coming third (though the 10-90\% range for the models
includes the possibility of it being the most favored).  This is not
surprising, given the single-peaked nature of the lower-mass sample.
The log normal and exponential models are likely disfavored due to
their long tails.  The multiple-bin histogram models are disfavored
because the structure of the lower-mass mass distribution is not
complicated enough to warrant the extra parameter space implied by the
larger number of bins.  The two Gaussian model is disfavored because
the distribution is single-peaked, so the second Gaussian is
superfluous.

@@

Another technique for computing $p(M_i|d)$ is the reversible-jump MCMC
\citep{Green1995}.  For a reversible-jump MCMC, we consider a
``super-model'' that encompasses all the models under consideration.
The parameter space of the super-model consists of a discrete
parameter that identifies the choice of model, $M_i$, and the
continuous parameters appropriate for this model, $\vtheta_i$.  We
denote a point in the super-model parameter space by $\{M_i,
\vtheta_i\}$; each such point is a statement that, e.g., ``the
underlying mass distribution is a Gaussian, with parameters $\mu$ and
$\sigma$,'' or ``the underlying mass distribution is a triple-bin
histogram with parameters $w_1$, $w_2$, $w_3$, and $w_4$,'' or ....
To compare models, we are interested in the quantity (see Equation
\eqref{eq:model-posterior-def})
\begin{equation}
  p(M_i|d) = \int d\vtheta_i p(M_i, \vtheta_i|d).
\end{equation}
If we perform an MCMC in the super-model parameter space, then we
obtain a chain of samples $\{M_i, \vtheta_i\}$ distributed in
parameter space with density $p(M_i,\vtheta_i) d\vtheta_i$ and we can
estimate the integral as
\begin{equation}
  p(M_i|d) = \int d\vtheta_i p(M_i, \vtheta_i|d) \approx \frac{N_i}{N},
\end{equation}
where $N_i$ is the number of MCMC samples that have discrete parameter
$M_i$, and $N$ is the total number of samples in the MCMC.  In other
words, the fraction of samples lying in the parameter space of model
$M_i$ gives the probability of that model relative to the other models
under consideration.

@@

The harmonic mean estimators for the model probabilities are displayed
in Figure \ref{fig:high-harmonic-evidence}; compare Figure
\ref{fig:harmonic-evidence}.  Reversible jump MCMC calculations of the
model probabilities are displayed in Figure
\ref{fig:high-rj-evidence}; compare Figure \ref{fig:rj}.  Though the
resulting rankings are not in complete agreement, a striking feature
of both calculations is that the two-Gaussian model is much more
likely with the higher-mass samples.  In the harmonic mean
calculation, the power-law model is the most likely; in the
reversible-jump calculation the exponential model is most likely.  In
either case, the most likely model differs significantly from the most
likely model for the lower-mass samples.  The model selection provides
more strong evidence that the wind-fed and Roche-lobe systems come
from different underlying populations.  The improvement of the
exponential model relative to the lower-mass analysis (it is the most
likely model in the reversible-jump MCMC) is encouraging for
theoretical calculations that attempt to model the entire population
of X-ray binaries with this mass model.  Note also that the increasing
structure of the mass distribution favors histogram models with three
bins over those with fewer bins.

\begin{figure}
  \begin{center}
    \plotone{plots/evidence-high}
  \end{center}
  \caption{\label{fig:high-harmonic-evidence} Harmonic mean evidence
    for the various models from Section \ref{sec:models} applied to
    all 20 systems in Table \ref{tab:sources} (see Section
    \ref{sec:direct-evidence}).  In increasing order along the
    $x$-axis, the models are the power-law of Section
    \ref{sec:power-law} (PL), the decaying exponential of Section
    \ref{sec:exponential} (E), the single Gaussian of Section
    \ref{sec:gaussian} (G), the double Gaussian of Section
    \ref{sec:gaussian} (TG), and the one-, two-, three-, four-, and
    five-bin histogram models of Section
    \ref{sec:non-parametric-models} (H1, H2, H3, H4, H5,
    respectively).  The errorbars are bootstrap estimates of the 10\%
    and 90\% bounds on the harmonic mean estimator.  Compare to Figure
    \ref{fig:harmonic-evidence}.}
\end{figure}

>>>>

\begin{abstract}
  We perform a Bayesian analysis of the mass distribution of
  stellar-mass black holes using the observed masses of 15 low-mass
  X-ray binary systems undergoing Roche lobe overflow and five
  higher-mass, wind-fed X-ray binary systems.  Using MCMC
  calculations, we model the mass distribution both
  parametrically---as a power law, exponential, gaussian, combination
  of two gaussians, or log-normal distribution---and
  non-parametrically---as histograms with varying numbers of bins.  We
  provide confidence bounds on the shape of the mass distribution in
  the context of each model and compare the models with each other by
  calculating their Bayesian evidence as supported by the
  measurements.  The mass distribution of the low-mass systems is best
  fit by a power-law, while the distribution of the complete sample is
  best fit by the exponential model.  We examine the existence of a
  ``gap'' between the most massive neutron stars and least massive
  black holes, finding that the best model (the power law) fitted to
  the low-mass systems only gives a minimum black-hole mass of 4.3
  $\Msun$ (90\% confidence), while the best model (the exponential)
  fitted to all 20 systems gives a minumum black-hole mass of 4.5 and
  5.2 $\Msun$ (90\% confidence), respectively.  We conclude that our
  sample of black hole masses provides strong evidence of a gap
  between the maximum neutron star mass and minimum black hole mass.
\end{abstract}

@@

The mass function defines a lower limit on the mass: $f(M) < M$.  To
accurately determine the mass of the black hole, the inclination $i$
and mass ratio $q$ must be measured.  Ideally, this can be
accomplished by fitting ellipsoidal light curves and study of the
rotational broadening of spectral lines from the secondary, but even
in the most studied case (see, e.g., \citet{Cantrell2010} on A0620)
this procedure is complicated.  In particular, contributions from an
accretion disk and hot spots in the disk can significantly distort the
measured inclination and mass ratios.  For some systems (e.g.\ GS 1354
\citep{Casares2009}) strong variability completely prevents
determination of the inclination from the lightcurve; in these cases
an upper limit on the inclination often comes from the observed lack
of eclipses in the lightcurve.  In general, accurately determining $q$
and $i$ requires a careful system-by-system analysis.

@@

@@

In this section we apply a statistical analysis to various models for
the underlying mass distribution from which the low-mass systems in
Table \ref{tab:sources} were drawn.  We will treat the combined low-
and higher-mass systems in Section \ref{sec:higher-mass}.  The end
result will be the probability distribution for the parameters of each
model implied by the data from Section \ref{sec:systems} in
combination with our prior assumptions about the probability
distribution for the parameters.  Bayes' rule relates these
quantities.  For a model with parameters $\vtheta$ in the presence of
data $d$, Bayes' rule states
\begin{equation}
  \label{eq:Bayes-rule}
  p(\vtheta | d) = \frac{p(d | \vtheta) p(\vtheta)}{p(d)}.
\end{equation}
Here, $p(\vtheta|d)$, called the posterior probability distribution
function, is the probability distribution for the parameters $\vtheta$
implied by the data $d$; $p(d|\vtheta)$, called the likelihood, is the
probability of observing data $d$ given that the model parameters are
$\vtheta$; $p(\vtheta)$, called the prior, reflects our estimate of the
probability of the various model parameters in the absence of any
data; and $p(d)$, called the evidence, is an overall normalizing
constant ensuring that
\begin{equation}
  \int d\theta\, p(\vtheta|d) = 1,
\end{equation}
whence
\begin{equation}
  \label{eq:evidence-def}
  p(d) = \int d\vtheta\, p(d|\vtheta) p(\vtheta).
\end{equation}

@@

Our calculation of the likelihood of each observation does not include
any attempt to account for selection effects in the observations.  We
simply assume (almost certainly incorrectly) that any black hole drawn
from the underlying mass distribution is equally likely to be
observed.  (See also \citet{Ozel2010}.)

@@

The marginalized distribution for $\alpha$ is broad, with
\begin{equation}
  -11.8 < \alpha < 6.8
\end{equation}
enclosing 90\% of the probability (excluding 5\% on each side).  We
have $p(\alpha < 0) = 0.6$.  The median value is $\alpha = -3.35$.
The broadness of the marginalized distribution for $\alpha$ comes from
the need to match the relatively narrow range in mass of the
lower-mass systems.  When $\alpha$ is negative, the resulting mass
distribution slopes down; $\Mmin$ is constrained to be near the lowest
mass of the observed black holes, while $\Mmax$ is essentially
irrelevant.  Conversely, when $\alpha$ is positive and the mass
distribution slopes up, $\Mmax$ must be close to the largest mass
observed, while $\Mmin$ is essentially irrelevant.  Figure
\ref{fig:power-law-2D} illustrates this effect, showing the
correlations between $\alpha$ and $\Mmin$ and $\alpha$ and $\Mmax$.
When we include the higher-mass systems in the analysis, the long tail
will eliminate this effect by bringing both $\Mmin$ and $\Mmax$ into
play for all values of $\alpha$.

@@

Figure \ref{fig:two-gaussian} shows the marginalized distributions for
the two-Gaussian model parameters from our MCMC runs.  We find $\alpha
> 0.8$ with 62\% probability, clearly favoring the first Gaussian.
The distribution for $\mu_1$ and $\sigma_1$ are similar to those of
the single Gaussian displayed in Figure \ref{fig:gaussian}, indicating
that the first Gaussian is centered around the peaks of the lower-mass
distributions.  The second Gaussian's parameter distributions are much
broader.  The second Gaussian appears to be sampling the tail of the
mass samples.  In spite of the ability to match a more complicated
distribution, we find that this model is strongly disfavored relative
to the single-Gaussian model for this dataset:
$p(\textnormal{Gaussian}|d) / p(\textnormal{Two Gaussian}|d) \simeq
4.7$ (see Section \ref{sec:model-selection} for discussion).

\begin{figure}
  \begin{center}
    \plotone{plots/two-gaussian}
  \end{center}
  \caption{\label{fig:two-gaussian} The marginal distributions for the
    five parameters of the two-Gaussian model.  We have $\alpha > 0.8$
    with 62\% probability, favoring the first of the two Gaussians.
    The distributions for $\mu_1$ and $\sigma_1$ are similar to those
    of the single Gaussian model displayed in Figure
    \ref{fig:gaussian}; the second Gaussian's parameter distributions
    are much broader (recall that we constrain $\mu_2 > \mu_1$).  The
    second Gaussian is attempting to fit the tail of the mass samples.
    The extra degrees of freedom in the distribution from the second
    Gaussian do not provide enough extra fitting power to compensate
    for the increase in parmeter space, however: the two-Gaussian
    model is disfavored relative to the single Gaussian by a factor of
    $4.7$ on this dataset (see Section \ref{sec:model-selection} for
    discussion).}
\end{figure}

@@

@@

A powerful technique for computing $p(M_i|d)$ is the reversible-jump
MCMC \citep{Green1995}.  Consider the ``super-model'' that encompasses
all the models under consideration.  The parameter space of the
super-model consists of a discrete parameter that identifies the
choice of model, $M_i$, and the continuous parameters appropriate for
this model, $\vtheta_i$.  We denote a point in the super-model
parameter space by $\{M_i, \vtheta_i\}$; each such point is a
statement that, e.g., ``the underlying mass distribution is a
Gaussian, with parameters $\mu$ and $\sigma$,'' or ``the underlying
mass distribution is a triple-bin histogram with parameters $w_1$,
$w_2$, $w_3$, and $w_4$,'' or ....  To compare models, we are
interested in the quantity (see Equation
\eqref{eq:model-posterior-def})
\begin{equation}
  p(M_i|d) = \int d\vtheta_i p(M_i, \vtheta_i|d).
\end{equation}
If we perform an MCMC in the super-model parameter space, then we
obtain a chain of samples $\{M_i, \vtheta_i\}$ distributed in
parameter space with density $p(M_i,\vtheta_i) d\vtheta_i$ and we can
estimate the integral as
\begin{equation}
  p(M_i|d) = \int d\vtheta_i p(M_i, \vtheta_i|d) \approx \frac{N_i}{N},
\end{equation}
where $N_i$ is the number of MCMC samples that have discrete parameter
$M_i$, and $N$ is the total number of samples in the MCMC.  In other
words, the fraction of samples lying in the parameter space of model
$M_i$ gives the probability of that model relative to the other models
under consideration.

@@

Reversible jump MCMC calculations of the model probabilities are
displayed in Figure \ref{fig:high-rj-evidence}; compare Figure
\ref{fig:rj}.  The exponential model is the most favored model for the
complete sample, with the two-Gaussian model the second-most favored.
The ranking of models differs significantly from the lower-mass
samples.  The model selection provides more strong evidence that the
wind-fed and Roche-lobe systems come from different underlying
populations.  The improvement of the exponential model relative to the
lower-mass analysis is encouraging for theoretical calculations that
attempt to model the entire population of X-ray binaries with this
mass model.  Note also that the increasing structure of the mass
distribution favors histogram models with three bins over those with
fewer bins.
